Childhood exposure to non-persistent endocrine disruptors, glucocorticosteroids, and attentional function: A study based on the parametric g-formula

### 2024-03-15 output/paper/manuscript.qmd
Re-writing discussion basically.

Our results are consistent with prior epidemiological research that associated exposure to certain non-persistent {{< acr edcs >}} with higher levels of glucocorticoids [@KimLeeMoon:2018; @SunLiJin:2018; @SearsLiuLanphear:2023]. Repeated measures up to 15 months of age of the phthalate metabolites MEHHP, MEOHP, MiBP, and MnBP showed positive associations with free cortisol in Korean children, with no effect modification by sex [@KimLeeMoon:2018]. In a cohort of Chinese pregnant women, phthalate metabolites were measured at 14, 24, and 36 weeks of gestation, and the glucocorticoids cortisol and cortisone were measured in cord blood. Third-trimester levels of MEHP were positively associated with cortisol, while MECPP and MEOHP were negatively associated with cortisone [@SunLiJin:2018]. Time- and chemical-dependent sex differences were also found: during the third trimester, MEHHP and MEOHP were positively associated with cortisol in females, while negatively associated in males [@SunLiJin:2018]. In a longitudinal study, a mixture of several phthalate metabolites, driven by MEP, MiBP, and MBzP, measured in childhood, showed a positive association with hair cortisol measured at 12 years [@SearsLiuLanphear:2023].

There are some differences, though. First, these studies only focus on phthalates, either as individual metabolites or as mixtures. Second, exposure assessment in @SunLiJin:2018 and @KimLeeMoon:2018 was performed during gestation or the first 15 months of life, respectively. Finally, the glucocorticosteroids were measured in cord blood [@SunLiJin:2018] and hair [@SearsLiuLanphear:2023]. Contrary to these studies, we did find effect modification by sex. We are not aware of other epidemiological studies investigating phthalates metabolites, phenols, and {{< acr opps >}}, in relation to urinary glucocorticosteroids in childhood. Nonetheless, previous toxicological studies provide evidence for the inhibition by phthalates of human 11$\beta$-hydroxysteroid dehydrogenase 2 (11$\beta$-HSD2) activity, responsible for the conversion of active cortisol into inactive cortisone [@ZhaoChuHuang:2010; @MaLianDong:2011]. There is also *in silico* evidence suggesting that {{< acr bpa >}}, a phenol, and Triazophos (TAP), a organophosphorus insecticide, can bind to the human glucocorticoid receptor [@PrasanthDivyaSadasivan:2010; @YangLiRen:2019].

>>>>

Prior epidemiological research provides preliminary evidence for an association between certain non-persistent {{< acr edcs >}} with higher levels of glucocorticoids [@KimLeeMoon:2018; @SunLiJin:2018; @SearsLiuLanphear:2023]. Repeated measures up to 15 months of age of the phthalate metabolites MEHHP, MEOHP, MiBP, and MnBP showed positive associations with free cortisol in Korean children, with no effect modification by sex [@KimLeeMoon:2018]. In a cohort of Chinese pregnant women, phthalate metabolites were measured at 14, 24, and 36 weeks of gestation, and the glucocorticoids cortisol and cortisone were measured in cord blood. Third-trimester levels of MEHP were positively associated with cortisol, while MECPP and MEOHP were negatively associated with cortisone [@SunLiJin:2018]. Time- and chemical-dependent sex differences were also found: during the third trimester, MEHHP and MEOHP were positively associated with cortisol in females, while negatively associated in males [@SunLiJin:2018]. In a longitudinal study, a mixture of several phthalate metabolites, driven by MEP, MiBP, and MBzP, measured in childhood, showed a positive association with hair cortisol measured at 12 years of age [@SearsLiuLanphear:2023].

Our results...
There are some differences, though.
First, these studies only focus on phthalates, either as individual metabolites or as mixtures.
Second, exposure assessment in @SunLiJin:2018 and @KimLeeMoon:2018 was performed during gestation or the first 15 months of life, respectively.
Finally, the glucocorticosteroids were measured in cord blood [@SunLiJin:2018] and hair [@SearsLiuLanphear:2023].

Contrary to these studies, we did find effect modification by sex.

We are not aware of other epidemiological studies investigating phthalates metabolites, phenols, and {{< acr opps >}}, in relation to urinary glucocorticosteroids in childhood.
Nonetheless, previous toxicological studies provide evidence for the inhibition by phthalates of human 11$\beta$-hydroxysteroid dehydrogenase 2 (11$\beta$-HSD2) activity, responsible for the conversion of active cortisol into inactive cortisone [@ZhaoChuHuang:2010; @MaLianDong:2011].
There is also *in silico* evidence suggesting that {{< acr bpa >}}, a phenol, and Triazophos (TAP), a organophosphorus insecticide, can bind to the human glucocorticoid receptor [@PrasanthDivyaSadasivan:2010; @YangLiRen:2019].
